### 2012-06-15 docs/Jesse/Proposal/BurgersNotes.tex
put together rough proposal with a few missing sections

\section{The viscous Burgers equation}

We will begin with the viscous Burgers' equation as an example of

\begin{figure}[!h]
\centering
\includegraphics[scale=.5]{figs/burgers1e4.png}
\caption{Shock solution for Burgers' equation with $\epsilon = 10^{-4}$.} 
\end{figure}