### 2012-06-15 docs/Jesse/Proposal/proposalIntro.tex

\section{Motivations}

Over the last three decades, the use of Computational Fluid Dynamics as an engineering and design tool in design of aircraft has become common.  Wind tunnel experiments are often complemented by computational simulations, and the simulation of off-design conditions.  Great advances have been made in the field since its introduction, especially in areas of meshing, computer architecture, and solution strategies.  Despite this, there still exist many computational limitations in existing CFD methods:  
\begin{itemize}
\item \textbf{Higher order methods :} Higher order methods stand to offer large computational savings through a more efficient use of discrete degrees of freedom.  However, there are very few working higher-order CFD codes in existence, and most higher order methods tend to degrade to first-order accuracy near shocks. The use of higher order codes to solve the steady state equations is even rarer, where convergence of discrete nonlinear steady equations is a tricky issue \cite{BoeingHigherOrder}.  
\item \textbf{Automatic adaptivity :} The use of adaptive meshes is crucial to many CFD applications, where the solution often exhibits very localized sharp gradients and shocks.  Good resolution for such problems under uniform meshes is computationally prohibitive and impractical for most physical regimes of interest.  However, the construction of ``good" meshes is a difficult task, usually requiring a-priori knowledge of the form of the solution.  An alternative to such is the construction of automatically adaptive schemes; such methods usually begin with a coarse mesh and refine based on the minimization of some error.  However, this task is difficult, as the convergence of numerical methods for problems in CFD is notoriously sensitive to mesh quality.  Additionally, the use of adaptivity becomes even more difficult in the context of higher order and $hp$ methods \cite{BoeingHigherOrder}.  
\end{itemize}
Both of these issues are tied to the notion of \emph{robustness}.  We define robustness loosely as the degradation of solution quality with respect

typical physical conditions of interest for the compressible Navier-Stokes equations involve low viscosity/high Reynolds numbers.

Higher order adaptivity and stability for CFD problems are issues for high Reynolds and Mach numbers. 
\todo{Address the fact that CFD problems have issues of robustness; introduce reasoning that we will study the model problem of convection-diffusion first to better understand this. }

\subsection{Singular perturbation problems and robustness}

Historically, the Galerkin method has been very successfully applied to a broad range of problems in solid mechanics, for which the variational problems resulting from the PDE are symmetric and coercive (positive-definite). It is well known that the finite element method produces optimal or near-optimal results for such problems, with the finite element solution matching or coming close to the best approximation of the solution in the finite element space. However, standard Bubnov-Galerkin methods tend to perform poorly for the class of PDEs known as singular perturbation problems. These problems are often characterized by a parameter that may be either very small or very large in the context of physical problems.  An additional complication of singular perturbation problems is that very often, in the limiting case of the parameter blowing up or decreasing to zero, the PDE itself will change types (e.g.\ from elliptic to hyperbolic).

A canonical example of a singularly perturbed problem is the convection-diffusion equation on domain $\Omega \in \mathbb{R}^3$,
\[
\div \left(\beta u\right) - \epsilon \Delta u = f.
\]
The equation represents the change in the scalar quantity $u$, representing the concentration of a quantity in a given medium, taking into account both convective and diffusive effects. $\beta \in \mathbb{R}^3$ specifies the direction and magnitude of convection, while the singular perturbation parameter $\epsilon$ represents the diffusivity of the medium. In the limit of an inviscid medium as $\epsilon\rightarrow 0$, the equation changes types, from elliptic to hyperbolic, and from second order to first order.

We will illustrate the issues associated with numerical methods for this equation using one dimensional examples.  In 1D, the convection-diffusion equation is
\begin{align*}
\beta u'-\epsilon u'' &= f.
\end{align*}
For Dirichlet boundary conditions $u(0)=u_0$ and $u(1)= u_1$, the solution can develop sharp boundary layers of width $\epsilon$ near the outflow boundary $x=1$.

\begin{figure}[!h]
\centering
\includegraphics[scale=.4]{figs/GalerkinOscTight.png}
\caption{Oscillations in the 1D finite element solution of the convection-diffusion equation for small diffusion.}
\label{fig:GalerkinOsc}
\end{figure}

The poor performance of the finite element method for this problem is reflected in the bound on the error in the finite element solution --- under the standard Bubnov-Galerkin method with $u\in H^1(0,1)$, we have the bound given in \cite{roos2008robust}:
\[
\|u-u_h\|_\epsilon \leq C \inf_{w_h}\|u-w_h\|_{H^1(0,1)},
\]
for $\|u\|_\epsilon^2 \coloneqq \|u\|_{L^2}^2 + \epsilon \|u'\|_{L^2}^2$, with $C$ independent of $\epsilon$. An alternative formulation of the above bound is 
\[
\|u-u_h\|_{H^1(0,1)} \leq C(\epsilon) \inf_{w_h}\|u-w_h\|_{H^1(0,1)},
\]
where $C(\epsilon)$ grows as $\epsilon\rightarrow 0$. The dependence of the constant $C$ on $\epsilon$ is referred to as a \textit{loss of robustness} --- as the singular perturbation parameter $\epsilon$ decreases, our finite element error is bounded more and more loosely by the best approximation error.  As a consequence, the finite element solution can diverge significantly from the best finite element approximation of the solution for very small values of $\epsilon$.  For example, Figure~\ref{fig:GalerkinOsc} shows an example of how, on a coarse mesh, and for small values of $\epsilon$, the Galerkin approximation of the solution to the convection-diffusion equation with a boundary layer develops spurious oscillations everywhere in the domain, even where the best approximation error is small.  These oscillations grow in magnitude as $\epsilon \rightarrow 0$, eventually polluting the entire solution.

\section{Goal}

To develop a robust, stable $hp$-adaptive scheme for the steady compressible laminar Navier-Stokes equations in transonic/supersonic regimes.
\todo{FINISH}

\section{Literature review}

For the past half-century, problems in CFD have been solved using a multitude of methods, many of which are physically motivated, and thus applicable only to a small number of problems and geometries. We consider more general methods, whose framework is applicable to different problems in CFD; however, our specific focus will be on the nonlinear shock problems of compressible aerodynamics. Broadly speaking, the most popular general methods include (in historical order) finite difference methods, finite volume methods, and finite element methods.

\subsection{Finite difference and finite volume methods}

For linear problems, finite difference (FD) methods approximate derivatives based on interpolation of pointwise values of a function.  FD methods were popularized first by Lax, who introduced the concepts of the monotone scheme and numerical flux. For the conservation laws governing compressible aerodynamics, FD methods approximated the conservation law, using some numerical flux to reconstruct approximations to the derivative at a point. Finite volume (FV) methods are similar to finite difference methods, but approximate the integral version of a conservation law as opposed to the differential form. FD and FV have roughly the same computational cost/complexity; however, the advantage of FV methods over FD is that FV methods can be used on a much larger class of discretizations than FD methods, which require uniform or smooth meshes.

For nonlinear shock problems, the solution often exhibits sharp gradients or discontinuities, around which the solution would develop spurious Gibbs-type oscillations. Several ideas were introduced to deal with oscillations in the solution near a sharp gradient or shock: artificial diffusion, total variation diminishing (TVD) schemes, and slope limiters. However, each method had its drawback, either in terms of loss of accuracy, dimensional limitations, or problem-specific parameters to be tuned \cite{Shu:Lectures}. Harten, Enquist, Osher and Chakravarthy introduced the essentially non-oscillatory (ENO) scheme in 1987 \cite{ENO}, which was improved upon with the weighted essentially non-oscillatory (WENO) scheme in \cite{WENO}. WENO remains a popular choice today for both finite volume and finite difference schemes. Most of these methods can be interpreted as adding some specific artificial diffusion to the given numerical scheme.  We refer to such schemes as \emph{modified equation} methods, as the exact solution no longer satisfies the discrete system due to the presence of additional artificial diffusion.

Historically, finite volumes and finite difference methods have been the numerical discretizations of choice for CFD applications; the simplicity of implementation of the finite difference method allows for quick turnaround time, and the finite volume method is appealing due to its locally conservative nature and flexibility. More recently, the finite element (FE) method has gained popularity as a discretization method for CFD applications for its stability properties and rigorous mathematical foundations. Early pioneers of the finite element method for CFD included Zienkiewicz, Oden, Karniadakis, and Hughes \cite{ChungCFDBook}.

\subsection{Stabilized finite element methods}

The finite element/Galerkin method has been widely utilized in engineering to solve partial differential equations governing the behavior of physical phenomena in engineering problems.  The method relates the solution of a partial differential equation (PDE) to the solution of a corresponding variational problem. The finite element method itself provides several advantages --- a framework for systematic mathematical analysis of the behavior of the method, weaker regularity constraints on the solution than implied by the strong form of the equations, and applicability to very general physical domains and geometries for arbitrary orders of approximation.

Traditionally, instability/loss of robustness in finite element methods has been dealt with using residual-based stabilization techniques.  Given some variational form, the problem is modified by adding to the bilinear form the strong form of the residual, weighted by a test function and scaled by a stabilization constant $\tau$.  The most well-known example of this technique is the streamline-upwind Petrov-Galerkin (SUPG) method, which is a stabilized FE method for solving the convection-diffusion equation using piecewise linear continuous finite elements \cite{SUPG}.  SUPG stabilization not only removes the spurious oscillations from the finite element solution of the convection-diffusion equation, but delivers the best finite element approximation in the $H^1$ norm.

\subsubsection{SUPG}

The Streamline Upwind Petrov Galerkin (SUPG) method is a stabilization method for $H^1$-conforming finite elements, the idea of which was originally motivated by artificial diffusion techniques in finite differences.  In particular, for the homogeneous 1D convection-diffusion equation, it is possible to recover, under a finite difference method, the exact solution at nodal points by adding an ``exact" artificial diffusion based on the mesh size $h$ and the magnitudes of the convection $\beta$ and the viscosity $\epsilon$.

The idea of ``exact" artificial viscosity was adapted to finite elements not through the direct modification of the equations, but through the test functions and weighting of the residual.  In particular, finite element and Galerkin methods are often referred to as ``weighted residual" methods, since the starting point of both is to multiply the residual by a particular test, or weighting, function.  Standard Galerkin methods, known as Bubnov-Galerkin methods, simply choose these weighting functions to be the same as the the basis functions used to approximate the solution.  A Petrov-Galerkin method refers to any Galerkin method other than a classical Bubnov-Galerkin method.

We will introduce the SUPG method at an abstract level for illustrative purposes only.  Further details and perspectives on the SUPG method can be found in \cite{SUPG}, as well as in an upcoming book by Hughes.  The convection-diffusion equation can be written as follows:
\[
Lu = \left(L_{\rm adv} + L_{\rm diff} \right) u = f,
\]
where $L_{\rm adv}u \coloneqq \div \left(\beta u\right)$ is the first order advective operator, and $L_{\rm diff}u \coloneqq \epsilon \Delta u$ is the second-order diffusive operator (then, roughly speaking, $L_{\rm diff} u = 0$).  Let us assume $u$ to be a linear combination of piecewise-linear basis functions $\phi_i, i = 0,\ldots,N$.  The SUPG method is then to solve
\[
\int_\Omega L u v + \int_\Omega \tau \left(L_{\rm adv}v\right)\left( L u - f\right) = \int_\Omega f v 
\]
for where $\tau$ is the SUPG parameter.  For uniform meshes in 1D, $\tau$ is chosen such that, for $f=0$, the matrix system resulting from SUPG is exactly equal to the finite difference system under ``exact" artificial diffusion.  However, unlike the artificial diffusion method, for $f\neq 0$, the SUPG method still delivers optimal stabilization.  In fact, the SUPG finite element solution in 1D is nothing less than the nodal interpolant and the best $H^1$ approximation of the exact solution, as seen in Figure~\ref{fig:SUPG}.

\begin{figure}[!h]
\centering
\subfigure[SUPG and standard FEM solutions]{\includegraphics[scale=.43]{figs/SUPG.png}}
\subfigure[SUPG test function]{\includegraphics[scale=.22]{figs/SUPGtest.png}}
\caption{SUPG and standard Bubnov-Galerkin solutions to the 1D convection-dominated diffusion equation, and a modified SUPG test function (in black) corresponding to a linear basis ``hat" function (in red).  The upwind portion of the element is emphasized, while the downwind portion is decreased.  The magnitude of the discontinuity between the upwind and downwind portion is controlled by the intrinsic timescale parameter $\tau$. }
\label{fig:SUPG}
\end{figure}

The idea of emphasizing the upwind portion of a test function is an old idea, introduced in 1977 by Zienkiewicz et al.\ in \cite{zienkUpwind}.  However, the precise amount of upwinding,\footnote{Insufficient upwinding results in a method which still exhibits oscillations and instabilities, while excessive upwinding leads to an overly diffusive method.} as well as the connection to residual-based stabilization methods, were novel to SUPG.

The method can be made to ``work" for $p>1$, but the results are not as strong as for linear elements, and there is no clear generalization of SUPG to higher order elements.  Additionally, the stabilization properties of SUPG are weaker in higher dimensions, and the SUPG solution is no longer the $H^1$ best approximation for 2D and 3D problems \todo{find citation}.  Despite these caveats, SUPG is still the most popular stabilization method of choice for convection-diffusion type problems, in both academic and industry applications.

An important feature of SUPG and other residual-based stabilization techniques that separates it from modified equation methods is the idea of \textit{consistency} --- by adding stabilization terms based on the residual, the exact solution still satisfies the same variational problem (i.e.\ Galerkin orthogonality still holds). Contrast this to the artificial diffusion methods in finite difference and finite volume methods, where a specific amount of additional viscosity is added based on the magnitude of the convection and diffusion parameters: unlike residual-based stabilization schemes, the exact solution to the original equation no longer satisfies the new stabilized formulation.

This addition of residual-based stabilization terms can be interpreted as a modification of the test functions as well.  For SUPG, the formulation can equivalently be written as
\[
\int_\Omega\left( Lu-f\right) v_i = \int_\Omega f v_i, \quad i = 0,\ldots,N
\]
where $v_i$ is defined as
\[
v_i = \phi_i(x) + \tau L_{\rm adv} \phi_i.  
\]
In other words, the test function $v_i$ is a perturbation of the basis function $\phi_i$ by a scaled advective operator applied to $\phi_i$.  For a linear $C^0$ basis function (the ``hat" function), this naturally leads to a bias in the upwind or streamline direction of the flow $\beta$, as seen in Figure~\ref{fig:SUPG}.

An important connection can now be made --- stabilization can be achieved by changing the test space for a given problem.  We will discuss in Section~\secref{optimalTest} approaching the idea of stabilization through the construction of \textit{optimal test functions} to achieve optimal approximation properties.

\subsubsection{DG methods}

Discontinuous Galerkin (DG) methods form a subclass of FEM; first introduced by Reed and Hill in \cite{Reed:73}, these methods were later analyzed by Cockburn and Shu \cite{CockburnShu:DG} and have rapidly gained popularity for CFD problems. Advantages of DG methods include the local conservation property, easily modified local orders of approximation, easy adaptivity in both $h$ and $p$, and efficient parallelizability. Rather than having a continuous basis where the basis function support spans several element cells, DG opts instead for a discontinuous, piecewise polynomial basis, where, like FV schemes, a \emph{numerical flux} facilitates communication between neighboring elements (unlike FV methods, however, there is no need for a reconstruction step).

The formal definition of the numerical flux (attributed to Peter Lax) on an element boundary is some function of the values on the edges of both the neighboring elements.  An additional reason for the popularity of DG methods is that they can be interpreted as stabilized FE methods through appropriate choices of this numerical flux \cite{Brezzi20063293}. We will illustrate this with the pure convection equation in 1D ---
\begin{align*}
\pd{\left(\beta u\right)}{x} &= f\\
u(0) &= u_0.
\end{align*}
The DG formulation is derived by multiplying by a test function $v$ with support only on a single element $K = [x_K,x_{K+1}]$ and integrating by parts. The boundary term is left alone, such that the local formulation is 
\[
\left.\beta u v\right|_{x_K}^{x_{K+1}} + \int_K -\beta u \pd{v}{x} = \int_K f v,
\]
and the global formulation is recovered by summing up all element-wise local formulations. However, the boundary term in the local formulation is presently ill-defined, as both $u$ and $v$ are dual-valued over element boundaries. Consequently, we make the choice to define the values of $u$ on the boundary (the \emph{traces} of $u$) as
\begin{align*}
u(x_K) \coloneqq u(x_K^-), \quad u(x_{K+1}) \coloneqq u(x_{K+1}^-),
\end{align*}
where $u(x_K^-)$ is the value of $u$ at $x_K$ as seen from the left, and $u(x_K^+)$ the value as seen from the right. Similarly, the traces of $v$ are defined to be
\begin{align*}
v(x_K) \coloneqq v(x_K^+), \quad v(x_{K+1}) \coloneqq v(x_{K+1}^-),
\end{align*}
We will refer to $v(x_K^+)$ as the upwind value of $v(x_K)$, and refer to DG under this specific choice of traces as upwind DG.  This specific choice of $v(x_K)$ as the upwind value is crucial; similarly to SUPG, the upwind DG emphasizes the test function in the direction of convection and changes the way the residual is measured.  As it turns out, the performance of DG for convection-type problems is closely tied to this upwinding --- choosing the value of $v(x_K)$ to be the downwind value $v(x_K^-)$ leads to an unstable method, while choosing $v(x_K)$ to be the average of the upwind and downwind values leads to a DG method with suboptimal stability properties, similar to an $H^1$-conforming continuous Galerkin approximation\cite{Brezzi20063293}.

Another perspective on the use of the numerical flux in DG methods is that the selection of specific DG fluxes imparts \emph{additional regularity} where needed.  For example, for the pure convection problem, the solution has a distributional derivative in the streamline direction, but is only $L^2$ in the crosswind direction. As a consequence of the regularity of the solution, the boundary trace of the solution is defined only in the direction of convection. The upwind DG method addresses the above issue by choosing the numerical flux to be the upwind flux; in this case, the DG numerical flux can be viewed as imparting additional regularity to the discrete solution than is implied by the continuous setting \cite{DPG1,DPG3}.

\subsubsection{HDG}

A more recent development in DG methods is the idea of \emph{hybridized} DG (HDG), introduced by Gopalakrishnan and Lazarov \cite{hybridDG}. The hybridized DG framework addresses several criticisms of common DG methods (globally coupled degrees of freedom, complicated/inefficient implementation procedures, suboptimal convergence of approximate fluxes) by identifying degrees of freedom with support only on element edges. Neighboring elements are coupled together only through these degrees of freedom, which can be interpreted as Lagrange multipliers enforcing weak continuity of the trial space. Additionally, HDG methods \todo{add bit about numerical fluxes still used, stabilization params}

\subsubsection{Optimization/NL solvers}

Reference (MIT, Fidkowski)

\todo{Find more nonlinear solvers}

\section{Scope}

This proposal will
\begin{enumerate}
\item Introduce DPG
\item Cover the problem of robustness for model problems
\item Extend results/formulations to nonlinear problems
\item Discuss proposed work
\end{enumerate}

\section{Range of CFD problems}

\begin{figure}[!h]
\centering
\includegraphics[scale=.45]{figs/CFD_tree.pdf}
\caption{A diagram of common CFD problems and their simplifying assumptions.}
\end{figure}

\begin{itemize}
\item Discuss full compressible Navier-Stokes with boundary layers
\item Discuss full compressible Navier-Stokes without boundary layers, connect to Euler (mention Dr Young and Boeing and the IBL connection \cite{BoeingDrela})
\item Discuss model problems
\end{itemize}