### 2012-06-15 docs/Jesse/Proposal/areaReq.tex
adding more figs/tex chapters to proposal

\section{Area requirements}

\subsection{Area A}

\begin{itemize}
\item{\textbf{Completed: }} Prove robustness of DPG method for the scalar convection-diffusion problem
\item Attempt analysis of the linearized Navier-Stokes system
\end{itemize}

\subsection{Area B}

\begin{itemize}
\item{\textbf{Completed: }} Collaborative work with Nathan Roberts on the higher order parallel adaptive DPG code Camellia.
\item Nonlinear DPG - Hessian adjoint trick
\item Adaptivity: anisotropic refinements and $hp$-schemes.
\item Distributed static condensation
\end{itemize}

\subsection{Area C}

\begin{itemize}
\item{\textbf{Completed: }} Convection-dominated diffusion, Burgers, and Navier-Stokes. 
\item Higher Reynolds number, ramp problem, Gaussian bump, airfoil.
\item Euler with NS regularization?
\end{itemize}