### 2012-07-24 docs/Jesse/PECOStalk/utdiss2-05/chapter-tables+figs.tex

\chapter{Making Tables and Including Figures}
\index{Making Tables and Including Figures@\emph{Making Tables
	and Including Figures}}%

The \emph{tabular} 
\index{commands!environments!tabular}%
environment allows us to create complex 
tables and figures, and draw boundaries around and within it.
The following example illustrates this:

\begin{table}[h]
\begin{center}
\caption{An example of a table.}
\vskip 10pt
\begin{tabular}{|ll|l|ll|l|lll|}
\cline{1-2} \cline{4-5} \cline{7-9}
\multicolumn{2}{|c|} {\textsl{Gegenwart}} & &
\multicolumn{2}{|c|} {\textsl{Imperfekt}} & &
\multicolumn{3}{|c|} {\textsl{Perfekt}} \\
\cline{1-2} \cline{4-5} \cline{7-9}
ich & bin  & & ich & war   & & ich & bin  & gewesen \\
du  & bist & & du  & warst & & du  & bist & gewesen \\
er  &      & & er  &       & & er  &      &         \\
sie & ist  & & sie & wart  & & sie & ist  & gewesen \\
es  &      & & es  &       & & es  &      &         \\
\cline{1-2} \cline{4-5} \cline{7-9}
wir & sind & & wir & waren & & wir & sind & gewesen \\
ihr & seid & & ihr & wart  & & ihr & seid & gewesen \\
sie & sind & & sie & waren & & sie & sind & gewesen \\
\cline{1-2} \cline{4-5} \cline{7-9}
Sie & sind & & Sie & waren & & Sie & sind & gewesen \\
\cline{1-2} \cline{4-5} \cline{7-9}
\end{tabular} \\[10pt]
Note: The assistance of Herr Professor Lothar Frommhold \\
in generating this table of German declensions \\
is gratefully acknowledged.
\vskip -20pt
\end{center}
\end{table}
\index{commands!environments!table}%

This table was created with the following sequence 
of commands:
\begin{verbatim}
\begin{table}[h]
\begin{center}
\caption{An example of a table.}
\vskip 10pt
\begin{tabular}{|ll|l|ll|l|lll|}
\cline{1-2} \cline{4-5} \cline{7-9}
\multicolumn{2}{|c|} {\textsl{Gegenwart}} & &
\multicolumn{2}{|c|} {\textsl{Imperfekt}} & &
\multicolumn{3}{|c|} {\textsl{Perfekt}} \\
\cline{1-2} \cline{4-5} \cline{7-9}
ich & bin  & & ich & war   & & ich & bin  & gewesen \\
du  & bist & & du  & warst & & du  & bist & gewesen \\
er  &      & & er  &       & & er  &      &         \\
sie & ist  & & sie & wart  & & sie & ist  & gewesen \\
es  &      & & es  &       & & es  &      &         \\
\cline{1-2} \cline{4-5} \cline{7-9}
wir & sind & & wir & waren & & wir & sind & gewesen \\
ihr & seid & & ihr & wart  & & ihr & seid & gewesen \\
sie & sind & & sie & waren & & sie & sind & gewesen \\
\cline{1-2} \cline{4-5} \cline{7-9}
Sie & sind & & Sie & waren & & Sie & sind & gewesen \\
\cline{1-2} \cline{4-5} \cline{7-9}
\end{tabular} \\[10pt]
Note: The assistance of Herr Professor Lothar Frommhold \\
in generating this table of German declensions \\
is gratefully acknowledged.
\vskip -20pt
\end{center}
\end{table}
\index{commands!environments!table}%
\end{verbatim}

The argument \texttt{h} indicates the position for the 
table, in this case ``here if possible''. Other values
of this argument are:
\texttt{t} (top of the page),
\texttt{b} (bottom of the page),
\texttt{p} (on the page of floats) and 
\texttt{H} (HERE! - requires using the package float.sty.
Note: When this option is used, LaTeX ignores all of its formatting
rules and does what you say, putting the entire float exactly where
it is defined. Check your output to make sure it is what you want!
If you are having trouble with LaTeX wanting to put a figure that's
larger than roughly half-a-page, as well as all of the figures
following it, at the end of a chapter, try using the command
\cn{clearpage} immediately following the large figure --- and maybe
a \cn{newpage} later.)
It is possible to combine several arguments, such as
\texttt{ht} (``here if possible, otherwise on top of
the page''). The default is \texttt{tbp}.

Figure \ref{f:ex} is a typical example of inclusion of a 
figure contained in an encapsulated PostScript file. 
\index{PostScript}%
\index{encapsulated PostScript}%
In order to use it, it is necessary to include the 
command \cn{usepackage\{psfig\}} 
\index{psfig}%
at the beginning of the document.

\begin{figure}[htb] % Imported eps example.
\begin{center}
\ \psfig{file=pup-on-rug.eps,height=1.5in,width=2.0in}
\caption{An example of an imported eps file.}
\label{f:ex}
\end{center}
\end{figure}
\index{commands!environments!figure}%
You can see the commands that generated this
figure in the source file. Look for the line
\cn{begin\{figure\}[htb] \% Imported eps example. }

The command that imports the file is \cn{psfig}, and it also 
controls its size (\texttt{height} and \texttt{width}), and 
can rotate the figure (\texttt{angle}).

Figures can also be drawn by using \LaTeX{} commands. 
Figure \ref{f:circuit} is an example 
(taken from \cite{gms:tlc}).

\begin{figure}[htb] % Picture example.
\begin{center}
   \setlength{\unitlength}{4mm}
   \begin{picture}(12,10)(-2,0)
      \linethickness{0.4pt}
      \qbezier(2.00,6.00)(7.00,6.00)(9.00,3.00)
      \qbezier(2.00,0.00)(7.00,0.00)(9.00,3.00)
      \qbezier(2.00,6.00)(4.00,3.00)(2.00,0.00)
      \qbezier(1.00,6.00)(3.00,3.00)(1.00,0.00)
      \put(9.75,3.00){\circle{1.50}}
      \put(10.50,3.00){\line(1,0){1.50}}
      \put(0.00,5.00){\line(1,0){1.50}}
      \put(0.00,1.00){\line(1,0){1.50}}
   \end{picture}
\caption{An example of a picture}
\label{f:circuit}
\end{center}
\end{figure}
\index{picture}%

The commands that generated this
picture are in the source file following the line
\cn{begin\{figure\}[htb] \% Picture example.  }

The commands used have rather obvious meanings. In particular, 
the command \cn{qbezier} 
\index{commands!qbezier@\cn{qbezier}}%
draws a quadratic Bezier curve, 
defined by its two ending points, and a third point (whose 
coordinates are in the middle) that is used as control point. 
Figure \ref{f:qb} illustrates the effect of the control point:

%\begin{figure}[htb] % Bezier curves example.
\begin{figure}[h] % Bezier curves example.
\begin{center}
   \setlength{\unitlength}{.8mm}
   \begin{picture}(55,55)(-15,0)
      \linethickness{1pt}
      \qbezier(0,0)(-10,30)(50,30)
      \qbezier(0,0)(20,50)(50,30)
      \thinlines
      \put(0,0){\line(-1,3){10}}
      \put(50,30){\line(-1,0){60}}
      \put(0,0){\line(2,5){20}}
      \put(50,30){\line(-3,2){30}}
      \put(0,0){\circle*{1}}
      \put(0,-1){\makebox(0,0)[t]{$A_{0,0}$}}
      \put(-10,30){\circle*{1}}
      \put(-10,31){\makebox(0,0)[b]{$B_{10,30}$}}
      \put(50,30){\circle*{1}}
      \put(58,29){\makebox(0,0)[b]{$C_{50,30}$}}
      \put(20,50){\circle*{1}}
      \put(20,51){\makebox(0,0)[b]{$D_{20,50}$}}
   \end{picture}
\caption{Bezier curves}
\label{f:qb}
\end{center}
\end{figure}
\index{Bezier curves}%

This figure has been generated with the following commands:
\begin{verbatim}
\begin{figure}[htb] % Bezier curves example.
\begin{center}
   \setlength{\unitlength}{.8mm}
   \begin{picture}(55,55)(-15,0)
      \linethickness{1pt}
      \qbezier(0,0)(-10,30)(50,30)
      \qbezier(0,0)(20,50)(50,30)
      \thinlines
      \put(0,0){\line(-1,3){10}}
      \put(50,30){\line(-1,0){60}}
      \put(0,0){\line(2,5){20}}
      \put(50,30){\line(-3,2){30}}
      \put(0,0){\circle*{1}}
      \put(0,-1){\makebox(0,0)[t]{$A_{0,0}$}}
      \put(-10,30){\circle*{1}}
      \put(-10,31){\makebox(0,0)[b]{$B_{10,30}$}}
      \put(50,30){\circle*{1}}
      \put(58,29){\makebox(0,0)[b]{$C_{50,30}$}}
      \put(20,50){\circle*{1}}
      \put(20,51){\makebox(0,0)[b]{$D_{20,50}$}}
   \end{picture}
\caption{Bezier curves}
\label{f:qb}
\end{center}
\end{figure}
\end{verbatim}