### 2013-04-29 docs/Jesse/SIAM CSE13/travelReport.tex
adding siam/rice talks

\documentclass[letterpaper,10pt]{article}

% last updated 1/14/07
\begin{document}
\author{Jesse Chan}
\title{Professional development travel report}
\maketitle

The SIAM CSE-13 conference is a conference held by the Society of Industrial and Applied Mathematics every two years, and seeks to enable in-depth discussions on a variety of computational efforts on problems in science and engineering.  The conference also seeks to foster the interdisciplinary culture required to meet large-scale challenges, and promote the training of the next generation of computational scientists.

Attending the 2013 SIAM CSE conference has benefited me professionally in several ways.  I attended a Student Career seminar, where I was able to ask four panelists (representing industry, academia, and the National Laboratories) several questions about careers in the computational sciences.  In particular, I was able to gain a better understanding of both the strengths and weaknesses of having a strongly interdisciplinary degree, as well as how to overcome those weaknesses in applications (particularly to academia and industry).

I was able to hold in-depth discussions with several previous graduate students on the current state of their industry and company.  In particular, I was able to gauge the effect of recent policy changes in the DOE (concerning austerity measures and budget cuts that affect research grants to new and innovative small businesses) that may affect several companies in the upcoming years.

Finally, I was able to form new connections and be exposed to a broader set of research topics at the National Laboratories by giving a talk during a minisymposium session.  The session was organized primarily by research staff at Sandia National Laboratories, and the talks given were focused largely on problems of transport, rather than my standard area of high-order finite element methods.  In this way, I was able to both converse with other researchers on the applicability of my own method to problems of interest at the national and industrial level.

The author would like to thank the CSEM Program for their Graduate School Professional Development Award that helped make this possible.

\end{document}