### 2013-07-30 docs/Jesse/Dissertation/acknowledgements.tex

This dissertation represents five years of work and study, which seem to have simply flown by much quicker than I expected.  First and foremost, I have to thank my advisor (and self-designated local parent figure during my time in Austin) Leszek Demkowicz.  Your insight, experience, and most of all, your drive to think outside the box were irreplaceable in building up both my work and my understanding throughout these last 5 years.  I have very much enjoyed all the fruitful discussions I have had with you, and have relished the fact that as I've mature mathematically, I've been able to contribute my own share into those discussions as well.  My experience in Austin would not have been nearly as pleasant without your kindness as well; your hospitality and warmth are something all of your students appreciate greatly.

None of the work in this thesis would have been possible without the help and expertise of fellow graduate student Nate Roberts, who almost singlehandedly built and maintains the parallel $hp$-adaptive DPG codebase Camellia, with which I ran all numerical experiments in this thesis.  Thanks for teaching me software design, for your help in debugging my code, for helpful discussions on each of our mathematical problems, and for good conversations and company in and outside of the ACES (now POB) building.

I'm grateful to the members of my committee -- Professors Robert Moser, Todd Arbogast, Omar Ghattas, and Venkat Raman -- for helpful discussions and feedback in evaluating my work, as well as for perspective on how my thesis work fits into the larger scheme of computational science and engineering.

To my co-authors on several papers, thank you for all your help in pushing me further along the path to this doctoral degree.  To Norbert Heuer, your mathematical insight has greatly advanced our understanding of the DPG method.  To Tan Bui, thanks for your emphasis on mathematical rigor, and for your incredible understanding of both application areas and numerical methods and their mathematics.  ICES is lucky to have you as a faculty member.  To Jay Gopalakrishnan, thanks for our communications, and for keeping me updated with your own insights on DPG.  Your interest and excitement in this area are contagious, and your students and colleagues all agree that you are one of the most kind and gracious professors they have ever worked with.  To John Evans, congratulations on your new position -- it was a pleasure to be able to write a paper with you, to hear your insights, and connect our two separate areas together.

For helpful discussions on both the mathematical and physical nature of my problem, I'm grateful to a host of folks at ICES.  To Ivo Babu\v{s}ka, thanks for allowing me to host the ICES Forum with you for 3 years, and for sharing your wealth of knowledge on $hp$ methods and singularities with me.  To Rhys Ulerich, thanks for discussions and answers to all of the strangest and most esoteric questions concerning compilers, linking, and all things code-based in general.  To Todd Oliver, thanks for the discussions on nondimensionalization and boundary conditions.  Roy Stogner and Paul Baumann thanks for giving me some of PECOS' vast perspective on common problems in fluid dynamics simulations using finite elements.  Paul Tsuji, thanks for the discussions on conditioning and Schur complement systems.  Truman Ellis and Jamie Bramwell, thanks for listening to me talk on DPG and mathematics -- when working on such a new topic, I am grateful for any colleagues who also understand it, with whom to share both successes and failures.

The Texas Advanced Computing Center (TACC) at The University of Texas at Austin has been very generous in providing HPC resources that have contributed to the research results reported within this paper, as well as consultation for whenever I experienced issues on their machines.  Thanks also to Chris Simmons and Karl Schulz for helping me figure out some of the tricks of the trade when it came to modern software engineering.

To my friends at ICES, thanks for making all the long hours spent in this building enjoyable.  Fred Qiu, it was great having you both as a roommate and colleague.  John Hawkins, thanks for your hospitality, and for the fun nights of bike building and conversation with you and Christa.  Chris Mirabito, Paul Tsuji, Henry Chang, Omar Al Hinai, Prapti Neupane, and the rest of the cribbage crowd -- may your fifteens be numerous.  Jeff Hussman, Jesse Kelly, it was fun to have someone with whom I could share both my math and my music.  Fred Nugen, it was great biking with you (you still owe me for those rims, by the way).  To so many folks at ICES -- Jessica, Lindley, Hamid, Vikram, Michele, Nick, James, and still so many more -- thanks for all the good conversation and camaraderie over these last few years.

To my community at Vox Veniae, you have been a pillar of support for the last 5 years of my life.  To list all the incredible people who have made an impact on my life would take almost the entire length of this acknowledgment section all over again.

Thanks to my family -- my parents Henry and Wendy, and my brother David -- thanks for the support throughout the last 5 years, and for always welcoming me home whenever I needed a break.  I love you all so much.

I would be remiss if I did not acknowledge the financial support I received during the course of this project.  This work was made possible by the CAM Fellowship at ICES and by
the Department of Energy [National Nuclear Security Administration] under Award Number [DE-FC52-08NA28615].