### 2013-07-30 docs/Jesse/Dissertation/dissPres.tex

\documentclass[compress]{beamer}
\mode<presentation>
\usetheme{Warsaw}
\usecolortheme{seagull}
%\useoutertheme[subsection=false]{smoothbars}
\useoutertheme{infolines}
\useinnertheme{rectangles}

\setbeamercovered{dynamic}

\usepackage{array}
\usepackage{amsmath,amssymb,amsfonts,mathrsfs,amsthm}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{mathtools}
\usepackage{dsfont}
\usepackage{pdfpages}
\usepackage[textsize=footnotesize,color=green]{todonotes}
\usepackage{algorithm, algorithmic}
\usepackage{bm}
\usepackage{tikz}
\usepackage[normalem]{ulem}

\usepackage{graphicx}
\usepackage{subfigure}
%\usepackage{caption}
%\usepackage{subcaption}

\usepackage{color}
\usepackage{undertilde}
\usepackage{pdflscape}
\usepackage{pifont}

\usepackage{bibentry}
\nobibliography*

%\usepackage[osf]{mathpazo}
%\usepackage{mathpazo}
%\renewcommand\rmdefault{ptm}

\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

\newcommand{\vect}[1]{\ensuremath\boldsymbol{#1}}
\newcommand{\tensor}[1]{\underline{\vect{#1}}}
\newcommand{\del}{\triangle}
\newcommand{\grad}{\nabla}
\newcommand{\curl}{\grad \times}
\renewcommand{\div}{\grad \cdot}
\newcommand{\ip}[1]{\left\langle #1 \right\rangle}
\newcommand{\eip}[1]{a\left( #1 \right)}
\newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
\newcommand{\pdd}[2]{\frac{\partial^2#1}{\partial#2^2}}

\newcommand{\circone}{\ding{192}}
\newcommand{\circtwo}{\ding{193}}
\newcommand{\circthree}{\ding{194}}
\newcommand{\circfour}{\ding{195}}
\newcommand{\circfive}{\ding{196}}

\newcommand{\Reyn}{\rm Re}

\newcommand{\bs}[1]{\boldsymbol{#1}}
\DeclareMathOperator{\diag}{diag}

\newcommand{\equaldef}{\stackrel{\mathrm{def}}{=}}

\newcommand{\tablab}[1]{\label{tab:#1}}
\newcommand{\tabref}[1]{Table~\ref{tab:#1}}

\newcommand{\theolab}[1]{\label{theo:#1}}
\newcommand{\theoref}[1]{\ref{theo:#1}}
\newcommand{\eqnlab}[1]{\label{eq:#1}}
\newcommand{\eqnref}[1]{\eqref{eq:#1}}
\newcommand{\seclab}[1]{\label{sec:#1}}
\newcommand{\secref}[1]{\ref{sec:#1}}
\newcommand{\lemlab}[1]{\label{lem:#1}}
\newcommand{\lemref}[1]{\ref{lem:#1}}

\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\nor}[1]{\left\| #1 \right\|}
\newcommand{\snor}[1]{\left| #1 \right|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
\newcommand{\LRc}[1]{\left\{ #1 \right\}}
\newcommand{\tanbui}[2]{\textcolor{blue}{\sout{#1}} \textcolor{red}{#2}}
\newcommand{\Grad} {\ensuremath{\nabla}}
\newcommand{\Div} {\ensuremath{\nabla\cdot}}
\newcommand{\Nel} {\ensuremath{{N^\text{el}}}}
\newcommand{\jump}[1] {\ensuremath{\LRs{\![#1]\!}}}
\newcommand{\uh}{\widehat{u}}
\newcommand{\fnh}{\widehat{f}_n}
\renewcommand{\L}{L^2\LRp{\Omega}}
\newcommand{\pO}{\partial\Omega}
\newcommand{\Gh}{\Gamma_h}
\newcommand{\Gm}{\Gamma_{-}}
\newcommand{\Gp}{\Gamma_{+}}
\newcommand{\Go}{\Gamma_0}
\newcommand{\Oh}{\Omega_h}

\newcommand{\eval}[2][\right]{\relax
  \ifx#1\right\relax \left.\fi#2#1\rvert}

\def\etal{{\it et al.~}}

\def\arr#1#2#3#4{\left[
\begin{array}{cc}
#1 & #2\\
#3 & #4\\
\end{array}
\right]}
\def\vecttwo#1#2{\left[
\begin{array}{c}
#1\\
#2\\
\end{array}
\right]}
\def\vectthree#1#2#3{\left[
\begin{array}{c}
#1\\
#2\\
#3\\
\end{array}
\right]}
\def\vectfour#1#2#3#4{\left[
\begin{array}{c}
#1\\
#2\\
#3\\
#4\\
\end{array}
\right]}

\newcommand{\G} {\Gamma}
\newcommand{\Gin} {\Gamma_{in}}
\newcommand{\Gout} {\Gamma_{out}}

% removes nav symbols
\beamertemplatenavigationsymbolsempty

% defines newblock as null, giving compile issues otherwise
\let\newblock\relax 
\institute[ICES]{Institute for Computational Engineering and Sciences}

\title[DPG for convection-diffusion]{A Discontinuous Petrov-Galerkin method for convection-diffusion problems}
\date{July 17, 2013}
\author[J. Chan, L. Demkowicz, R. Moser]{Jesse Chan \\ Supervisors: Leszek Demkowicz, Robert Moser \\ Committee: Todd Arbogast, Omar Ghattas, Venkat Raman}

\begin{document}
\begin{frame}
\maketitle
\end{frame}

\begin{frame}
\frametitle{Acknowledgements}
\begin{columns}[c]
\begin{column}{.03\textwidth}
\end{column}
\begin{column}{.33\textwidth}

Omar Al-Hinai

Ivo Babu\v{s}ka

Paul Baumann

Tan Bui

Henry Chang

Truman Ellis

John Evans

\end{column}
\begin{column}{.33\textwidth}

Jay Gopalakrishnan

Courtney Grymonprez

John Hawkins

Norbert Heuer

Jessica Meixner

Steve Mattis

Todd Oliver

\end{column}
\begin{column}{.33\textwidth}

Weifeng Qiu

Nate Roberts

Karl Schulz

Chris Simmons

Roy Stogner

Paul Tsuji

Rhys Ulerich

\end{column}

\end{columns}
\vspace{1cm}
\begin{center}
\ldots and many more.
\end{center}
\end{frame}

\section{Introduction/Literature review}
\frame{
\frametitle{Phenomena in compressible flow}
\begin{figure}
\centering
\subfigure[Shock wave]{\includegraphics[scale=.0704]{figs/compFlowPics/bullet_shock.png}}
\subfigure[Boundary layer]{\includegraphics[scale=.16]{figs/compFlowPics/boundary_layer.png}}
%\subfigure{\includegraphics[scale=.15]{figs/compFlowPics/capsule.png}}
\end{figure}
Compressible flow plays an important role in the aerospace and energy industries - transonic and supersonic aircraft, combustion engines, etc. 
}
\frame{
\frametitle{Compressible Navier-Stokes equations}
\begin{columns}[c]
\begin{column}{.475\textwidth}
Numerical difficulties:
\begin{itemize}
\item{} Resolving solution features (sharp, localized viscous-scale phenomena)
\begin{itemize}
\item{} Shocks
\item{} Boundary layers
\item{} Turbulence (non-localized)
\end{itemize}
\item{} Nonlinear convergence and uniqueness of solutions
\item{} Stability of numerical schemes
\begin{itemize}
\item{} Coarse/adaptive grids
\item{} Higher order
\end{itemize}
\end{itemize}
\end{column}
\begin{column}{.475\textwidth}
\begin{figure}
\centering
\includegraphics[scale = .7]{figs/compFlowPics/airfoil_mesh.png}
\end{figure}
\end{column}
\end{columns}
\vspace{.2cm}
\begin{center}
Idea: begin first with the model problem of convection-diffusion.
\end{center}
}

\begin{frame}{Outline}
  \tableofcontents[currentsection]
\end{frame}

\frame{
\frametitle{Robustness: convection-diffusion as a model problem}
\vspace{-.5cm}
\[
\div \left(\beta u\right) - \epsilon \Delta u = f, \quad \text{on }\Omega \in \mathbb{R}^3
\]
\vspace{-.5cm}
\begin{columns}[c]
\begin{column}{.49\textwidth}
In 1D: $\beta u' - \epsilon u'' = f$. Standard continuous Galerkin variational formulation: solve
\[
b(u,v) = \ell(v), \quad u,v\in H^1_0([0,1])
\]
where
\begin{align*}
b(u,v) &= \int_\Omega -\beta uv' + \epsilon u'v' \\
\ell(v) &= \int_\Omega f v 
\end{align*}
\end{column}
\begin{column}{.49\textwidth}
\begin{figure}
\centering
\includegraphics[scale=.35]{figs/GalerkinOscTight.png}
\caption{Solution for $f=1$. Oscillations in the Galerkin method for large Peclet numbers ${\rm Pe} \coloneqq \frac{h}{\epsilon}$.}
\end{figure}
\end{column}
\end{columns}
}

\frame{
\frametitle{Streamline-upwind Petrov-Galerkin (SUPG)}
SUPG solves $b_{\rm SUPG}(u,v) = l_{\rm SUPG}(v)$, where
\begin{align*}
b_{\rm SUPG}(u,v) &= b(u,v)+ \sum_{K} \int_{K} \tau \LRp{L_{\rm adv} v} L u \\
l_{\rm SUPG}(v) &= \ell(v) + \sum_{K} \int_{K} \tau \LRp{L_{\rm adv}v} f.
\end{align*}
\vspace{-.5cm}
\begin{columns}[c]
\begin{column}{.44\textwidth}
\vspace{-.5cm}
\begin{itemize}
\item{} $Lu = \div\left(\beta u\right) - \epsilon \triangle u$, $L_{\rm adv}u = \div \left(\beta u\right)$, and $\tau$ is a parameter. 
%\item{} Recovers ``exact'' artificial diffusion for $f=0$.
\item{} Effective for $f\neq 0$. %``exact'' artificial diffusion.
\item{} \emph{Residual-based} stabilization. 
\end{itemize}
\end{column}
\begin{column}{.54\textwidth}
\begin{figure}
\centering
\includegraphics[scale=.38]{figs/SUPG.png}
%\caption{Exact, Bubnov-Galerkin, and SUPG solutions.}
\end{figure}
\end{column}
\end{columns}
}

\frame{
\begin{columns}
\begin{column}{.49\textwidth}
Can be interpreted as a \emph{Petrov-Galerkin} method,
\[
b\LRp{u,\tilde{v}_{i}} = \ell\LRp{\tilde{v}_{i}}, \quad \forall i = 1,\ldots,N-1,
\]
where the SUPG test function $\tilde{v}_{i}$ is defined elementwise as\footnotemark
\[
\tilde{v}_{i} = \phi_i(x) + \tau L_{\rm adv} \phi_i.  
\]
\end{column}
\begin{column}{.49\textwidth}
\begin{figure}
\centering
\subfigure{\includegraphics[scale=.16]{figs/SUPGtest.png}}
\caption{SUPG test function $v_i$.}
\end{figure}
\end{column}
\end{columns}
\footnotetext{\tiny\bibentry{SUPG}}
}

\section{The DPG Method}
\begin{frame}{Outline}
  \tableofcontents[currentsection]
\end{frame}

\frame{
\frametitle{DPG: a minimum residual method via optimal testing}
Given a trial space $U$ and Hilbert test space $V$,
\[
b(u,v) = \ell(v), \quad u\in U, \quad v\in V
\]
This is equivalent to the operator equation posed in $V'$
\[
Bu = \ell
\]
if we identify $B:U\rightarrow V'$ and $\ell\in V'$ such that 
\begin{align*}
\langle Bu,v\rangle_V &\coloneqq b(u,v), \quad u\in U, v\in V,\\
\langle \ell,v\rangle_V &\coloneqq \ell(v), \quad v\in V.
\end{align*}
We seek to minimize the dual functional over $U_h \subset U$
\[
J(u_h) = \frac{1}{2}\|Bu_h-\ell\|_{V'}^2 \coloneqq\frac{1}{2} \sup_{v\in V\setminus\{0\}} \frac{| b(u_h,v)-\ell(v)|^2}{\nor{v}_V^2}.
\]
}

\begin{frame}
Let $R_V: V \to V'$ be the Riesz map (an isometry):\
\[
\langle R_V v,\delta v\rangle_V \coloneqq(v, \delta v)_V, \quad \forall \delta v \in V.
\]
Then, our functional $J(u_h)$ is equal to
\begin{equation*}
\min_{u_h\in U_h} J(u_h) = \frac{1}{2}\left\|Bu_h-\ell\right\|_{V'}^2 =  \frac{1}{2}\left\|R_V^{-1}(Bu_h-\ell)\right\|_V^2.
\end{equation*}
First order optimality: G\^ateaux derivative is zero in all directions $\delta u \in U_h$
\begin{align*}
\left(R_V^{-1}(Bu_h-\ell),R_V^{-1}B\delta u\right)_V &= 0, \quad \forall \delta u \in U_h.\\
\rightarrow \LRa{(Bu_h-\ell),R_V^{-1}B\delta u} &= 0, \\
\rightarrow b(u_h,R_V^{-1}B\delta u) - \ell(R_V^{-1}B\delta u) &= 0
\end{align*}
Summary: select test functions \textcolor{red}{$v_{\delta u} \coloneqq R_V^{-1}B\delta u$} to \textcolor{red}{minimize residuals}.  
\end{frame}

%\frame{
%\frametitle{Summary: select test functions to minimize residuals}
%For $\delta u \in U_h$, define the \textcolor{red}{optimal test function} $v_{\delta u}$. 
%\begin{equation*}
%v_{\delta u} \coloneqq R_V^{-1}B\delta u.
%\end{equation*} 
%Then, the residual $J(u_h) = \frac{1}{2}\left\|Bu_h-\ell\right\|_{V'}^2$ is minimized by the solution of
%\[
%b\LRp{u_h,v_{\delta u}} = \ell\LRp{v_{\delta u}}, \quad \forall \delta u \in U_h.
%\]
%}

\frame{
\frametitle{Practical details of DPG}
Computation of $v_{\delta u} \coloneqq R_V^{-1}B\delta u$ is \textbf{global} and \textbf{infinite-dimensional}.
\begin{itemize}
\item By choosing a \textcolor{red}{broken} test space $V$ and \textcolor{red}{localizable} norm $\nor{v}_V^2 = \sum_K \nor{v}_{V(K)}^2$, test functions can be determined locally.
\item In practice, we use an \textcolor{red}{enriched space} $V_h \subset V$, where $\dim(V_h) > \dim(U_h)$ elementwise, and \textcolor{red}{optimal test functions} are approximated by computing $v_{\delta u} \coloneqq R_{V_h}^{-1}B\delta u \in V_h$ through\footnote{\tiny\bibentry{DPG2}}
\[
\LRp{v_{\delta u},\delta v}_V = b(\delta u,\delta v), \quad \delta u\in U_h, \quad \forall \delta v\in V_h
\]
Typically, if $U_h = \mathcal{P}^p(\mathbb{R}^n)$, $V_h = \mathcal{P}^{p+\triangle p}(\mathbb{R}^n)$, where $\triangle p \geq n$.\footnote{\tiny\bibentry{practicalDPG}} 
\end{itemize}
\let\thefootnote\relax\footnotetext{\tiny\bibentry{DPG2}}
}

\frame{
\frametitle{Properties of DPG}
DPG provides a symmetric positive-definite stiffness matrix. Let $\{\phi_j\}_{j=1}^m$ be a basis for $U_h$, and $\{v_i\}_{i=1}^n$ a basis for $V_h$, such that $n>m$. Then, for 
\begin{align*}
B_{ji} &= b(\phi_j,v_i),\\
l_i &= \ell(v_i),
\end{align*}
DPG solves the discrete system for degrees of freedom $u$
\begin{align*}
\LRp{B^T R_V^{-1}B} u  &= \LRp{B^T R_V^{-1}} l,
\end{align*}
where, under a localizable norm and discontinuous test functions, $R_V^{-1}$ is block diagonal.
}

\frame{
\frametitle{Properties of DPG}
Additional properties of DPG include\footnotemark
\begin{itemize}
\item DPG provides the best approximation in the \textcolor{red}{energy norm}
\[
\|u\|_E = \|Bu\|_{V'} = \sup_{\nor{v}_V=1} \left|b(u,v)\right|.
\]
\item The energy error is computable through the residual
\[
\nor{u-u_h}_E = \nor{B(u-u_h)}_{V'} = \nor{R_V^{-1}(\ell-Bu_h)}_V = \nor{e}_V
\]
where the \textcolor{red}{error representation function} $e$ is defined through $(e,\delta v)_V = \ell(v)-b(u_h,\delta v)$ for all $\delta v\in V$. 
\end{itemize}
\footnotetext{\tiny\bibentry{DPG3}}
}

\frame{
\frametitle{Ultra-weak formulation}
\vspace{-.75cm}
\begin{columns}
\begin{column}{.64\textwidth}
Given a first order system $Au = f$, we identify the \textbf{\emph{partition}} $\Oh$ and \textbf{\textcolor{red}{mesh skeleton}} \textcolor{red}{$\Gh$}.
\end{column}
\begin{column}{.34\textwidth}
%\[
%\Oh = \cup_{j=1}^\Nel K_j, \quad \Gh = \cup_{j=1}^\Nel \partial K_j
%\]
\begin{figure}[!h]
\centering
\includegraphics[scale = .155]{mesh_skel.pdf}
\end{figure}
\end{column}
\end{columns}
\vspace{-.5cm}
The ultra-weak formulation for $Au = f$ on $\Oh$ is
\[
b\left(\left(u, \widehat{u}\right),v\right) \coloneqq \sum_K\langle \widehat{u},v \rangle_{\partial K} + (u,A_h^*v)_{\Oh}= \LRp{f,v}_{\Oh}.
\]
Under proper assumptions, $\sum_K\langle \widehat{u},v \rangle_{\partial K} = \langle \widehat{u}, \jump{v} \rangle_{\Gh}$, with energy setting
\[
u\in L^2\LRp{\Oh} \equiv L^2(\Omega), \quad v\in V=D(A^*_h), \quad
\widehat{u}\in \gamma(D(A)),
\]
where $D(A_h^*)$ is the broken graph space of the formal adjoint $A_h^*$, and $\gamma(D(A))$ the trace space of the graph space of operator $A$.\footnote{\tiny\bibentry{analysisDPG}}
}

\frame{
\frametitle{The canonical ``graph'' test norm}
Recall $\nor{u}_E \coloneqq \sup_{v\in V\setminus\{0\}}\frac{b(u,v)}{\nor{v}_V}$. Under the ultra-weak formulation, the trial norm 
\[
\nor{\LRp{u,\widehat{u}}}_U^2 \coloneqq \|u\|^2_{\L} + \|\widehat{u}\|_{\dag}^2
\] 
generates a test norm \emph{equivalent} to 
\[
\nor{v}_{V_{\rm opt}}^2 \coloneqq \|A_h^*v\|_{\L}^2 +\left(\sup_{\widehat{u}} \frac{\LRa{ \widehat{u},\jump{v} }_{\Gh}}{\|\widehat{u}\|}\right)^2.
\]
This norm is not localizable, so we instead substitute an $L^2$ term for the jump terms, giving us the (equivalent) \textcolor{red}{graph norm}\footnote{\tiny\bibentry{Bui-ThanhDemkowiczGhattas11b}}
\[
\nor{v}_V^2 \coloneqq \|A_h^*v\|_{\L}^2 + \nor{v}_{L^2}^2.
\]
\let\thefootnote\relax\footnotetext{\tiny{\dag: This norm has a marvelous technical definition for which this footnote is too small to contain.}}
}

\frame{
\frametitle{Global properties of test spaces: $L^2$-best approximations}

Under the ultra-weak formulation for $Au = f$.  
\[
b(\LRp{u,\uh},v) = \LRa{\widehat{u},\jump{v}}_{\Gh} + (u,A_h^*v)_{\Oh}= \LRp{f,v}_{\Oh},
\]
we recover the \textcolor{red}{$L^2$-projection} for \textcolor{red}{conforming} test functions that satisfy the adjoint equation:
\[
A_h^*v = \delta u, \quad \forall \delta u \in U_h. %v\in V^r = \LRc{v \in V(\Oh): \LRa{\uh,\jump{v}}_{\Gh} = 0, \quad \forall \LRp{0,\widehat{\delta u}} \in U}.
\]

For $\nor{v}_{\rm graph}^2 = \nor{A_h^*v}_{\L}^2 + \delta\nor{v}_{\L}^2$ as $\delta \rightarrow 0$; test functions solve
\[
\LRp{A_h^*v,A_h^*\delta v}_{L^2(K)} = (u,A_h^*\delta v)_{L^2(K)}.
\]
What can we say about \textcolor{red}{global properties} of local test functions?
}

\frame{
\frametitle{Global properties of DPG test spaces}
If we define $V^r_{{\rm opt}}$ as the optimal test space constructed using $V^r_p$, the \textcolor{red}{weakly conforming} space $\LRc{v \in V, \LRa{\widehat{\delta u},\jump{v}}_{\Gh} = 0, \quad \forall \LRp{0,\widehat{\delta u}} \in U_h}$,
\[
V^r_{{\rm opt}} = \LRc{ R_{V^r_p}^{-1}B \delta u, \quad \forall \delta u \in U_h},
\]
we have the following theorem\footnote{\tiny\bibentry{globalLocalDPG}}
\begin{theorem}
$V^r_{{\rm opt}} \subset V_{{\rm opt}}$.  Furthermore, the \textbf{field solutions} $u$ of $b(\LRp{u,\uh},v)=l(v)$ over $V_{{\rm opt}}$ and ${V}^r_{\rm opt}$ coincide.
\end{theorem}
\begin{itemize}
\item Locally determined test functions approximate global test functions.
\item Under a variant of the graph norm, \textcolor{red}{DPG approximates test functions which deliver $L^2$ optimality}.
\end{itemize}
}

%\frame{
%\frametitle{DPG as a non-conforming DG method on $V$}
%For DPG using the ultra-weak variational formulation
%\[
%b\left(\left(u, \widehat{u}\right),v\right) \coloneqq \LRa{\widehat{u},\jump{v}}_{\Gh} + (u,A_h^*v)_{\Oh}= \LRp{f,v}_{\Oh},
%\]
%under the graph test norm $\nor{v}_{\rm graph} = \nor{A_h^*v}_{\L} + \delta\nor{v}_{\L}$, as $\delta\rightarrow 0$, weakly conforming optimal test functions $\tilde{v}\in \tilde{V}$ satisfy 
%\[
%\LRp{A_h^*\tilde{v},A_h^*\tilde{\delta v}}_{\L} = (u,A_h^*\tilde{\delta v})_{\Oh}
%\]
%which yields a global \textcolor{red}{non-conforming least squares} method for $A_h^*v = u$.
%%\[
%%b\left(\left(u, \widehat{u}\right),\tilde{v}\right) = (u,A_h^*v)_{\Oh}= \LRp{f,v}_{\Oh}.
%%\]
%}

%\frame{%}

\subsection{A Robust DPG Method for Convection-Diffusion}
\begin{frame}{Outline}
  \tableofcontents[currentsection]
\end{frame}

\frame{
\frametitle{Ultra-weak formulation for convection-diffusion}
\vspace{-.25cm}
\begin{columns}
\begin{column}{.67\textwidth}
Convection-diffusion as a first order system is
\[
A \LRp{u,\sigma} \coloneqq 
\LRs {
\begin{array}{c}
\div (\beta u - \sigma) \\ \frac{1}{\epsilon}\sigma - \grad u
\end{array}} = \LRs{
\begin{array}{c}
f \\ 0
\end{array}
}.
\]
\end{column}
\begin{column}{.32\textwidth}
\begin{figure}[!h]
\centering
\includegraphics[width = 1.25in]{figs/hybridDGnodes.pdf}
\end{figure}
\end{column}
\end{columns}
\vspace{-.25cm}
The variational formulation is
\begin{align*}
b\left(\left(u,\sigma, \widehat{u}, \widehat{f}_n\right),
\left( v, \tau \right)\right) = \left(u,\grad_h\cdot \tau - \beta \cdot \grad_h
v\right)_{\Oh} + \left(\sigma, \epsilon^{-1} \tau + \grad_h v\right)_{\Oh}&\\
  - \LRa{\jump{\tau\cdot n}, \widehat{u} }_{\Gh} + \LRa{\widehat{f}_n,\jump{v} }_{\Gh}&,
\end{align*}
where $\widehat{f}_n \coloneqq \beta_n u - \sigma_n$ and $\LRa{\widehat{f}_n,\jump{v} }_{\Gh}$ is defined 
\[
\LRa{\widehat{f}_n,\jump{v} }_{\Gh}\coloneqq \sum_K\int_{\partial K}{\rm sgn}(\vec{n}) \, \widehat{f}_n v.
\]
}

\frame{
\frametitle{Graph norm under convection-diffusion}
For convection-diffusion, the graph test norm is defined elementwise
\[
\|\left(v,\tau\right)\|_{V(K)}^2 = \| \div \tau - \beta \cdot \grad v
\|_{L^2(K)}^2 + \| \epsilon^{-1} \tau + \grad v \|_{L^2(K)}^2 +
\|v\|_{L^2(K)}^2.
\]
Problem with this test norm: approximability of test functions.\footnote{\tiny\bibentry{globalLocalDPG}}
\vspace{-.3cm}
\begin{figure}[!h]
\centering
\subfigure[Element $\tau$]{\includegraphics[width=1.1in]{figs/u_graph_tau_mag_K.pdf}}
\subfigure[Element $v$]{\includegraphics[width=1.1in]{figs/u_graph_v_K.pdf}}
\subfigure[Global $\tau$]{\includegraphics[width=1.1in]{figs/u_graph_tau_mag.pdf}}
\subfigure[Global $v$]{\includegraphics[width=1.1in]{figs/u_graph_v.pdf}}
\caption{$v$ and $\tau$ components of optimal test functions for $u = 1$ on $K$, for $\beta = (1,0)$ and $\epsilon = 0.01$. }
\label{fig:optTestBoundary}
\end{figure}
}

\frame{
\frametitle{Determining an alternative test norm}
Recall the convection-diffusion bilinear form
\begin{align*}
b\left(\left(u,\sigma, \widehat{u}, \widehat{f}_n\right),
\left( v, \tau \right)\right) = \left(u,\div \tau - \beta \cdot \grad
v\right)_{\Oh} + \left(\sigma, \epsilon^{-1} \tau + \grad v\right)_{\Oh}&\\
 - \LRa{\jump{\tau\cdot n}, \widehat{u} }_{\Gh} + \LRa{\widehat{f}_n,\jump{v} }_{\Gh}&,
\end{align*}
We recover $\nor{\LRp{u,\sigma}}^2_{L^2(\Omega)}$ by choosing conforming $\LRp{v,\tau}$ satisfying the \emph{adjoint equations}
\begin{align*}
\div\tau - \beta \cdot \grad v   &= u\\
\frac{1}{\epsilon}\tau + \grad v &= \sigma
\end{align*}
with boundary conditions s.t.\ $\LRa{\jump{\tau\cdot n}, \widehat{u} }_{\Gamma}$ and $\LRa{ \widehat{f}_n,\jump{v} }_{\Gamma}$ vanish.
}

\frame{
\frametitle{A robust bound}
``Necessary'' conditions for robustness --- let $\boldsymbol U = \LRp{u,\sigma,\widehat{u},\widehat{f}_n}$. Then, by choosing specific $\LRp{v,\tau}$ satisfying the adjoint equations,
\[
\nor{\LRp{u,\sigma}}^2_{L^2(\Omega)} = b\LRp{{\boldsymbol U},\LRp{v,\tau}} = \frac{b\LRp{{\boldsymbol U},\LRp{v,\tau}}}{\nor{\LRp{v,\tau}}_V} \nor{\LRp{v,\tau}}_V \leq \nor{\boldsymbol U}_E \nor{\LRp{v,\tau}}_V
\]
Let $\lesssim$ denote a robust bound - \textcolor{red}{if $ \nor{\LRp{v,\tau}}_V \lesssim \nor{\LRp{u,\sigma}}_{\L}$}, then we have that
\[
\nor{\LRp{u,\sigma}}_{\L} \lesssim \nor{\boldsymbol U}_E
\]
\textbf{Main idea: the \textcolor{red}{test norm} should measure \textcolor{red}{adjoint solutions} robustly.}
}

\frame{
\frametitle{Choice of inflow boundary condition}

\begin{columns}
\begin{column}{.49\textwidth}
We impose the standard outflow wall boundary condition on $u$. For inflow boundary condition: 
\begin{itemize}
\item{} The standard choice of inflow boundary condition: $u = u_0$.
\item{} We impose a new ``conserved flux'' inflow condition: $\widehat{f}_n \coloneqq \beta_n u - \sigma_n \approx \beta_n u_0$ on $\Gamma_{\rm in}$.
\end{itemize}
\end{column}

\begin{column}{.49\textwidth}
\begin{figure}
\includegraphics[scale=.23]{figs/primal.png}
\caption{Non-standard ``conserved flux'' inflow.}
\end{figure}
\end{column}
\end{columns}
When $\sigma_n \approx 0$ near the inflow (i.e.\ most problems of interest where $\epsilon \ll 1$), condition on $\widehat{f}_n$ approximates condition on $u$. 
}

\frame{
\frametitle{Dirichlet inflow boundary condition}
Standard choice of boundary condition: $u = u_0$ on inflow boundary $\Gamma_{\rm in}$, induces boundary layers in adjoint problems, \textcolor{red}{$\nor{\beta\cdot \grad v}_{L^2} = O(\epsilon^{-1})$}.
\begin{figure}[h!]
\centering
\subfigure[Primal problem]{\includegraphics[scale=.23]{figs/primalDir.png}}
\subfigure[Adjoint problem]{\includegraphics[scale=.23]{figs/adjointDir.png}}
\caption{For the standard Dirichlet inflow condition, the solution to the adjoint problem can develop strong boundary layers at the outflow of the adjoint problem. }
\end{figure}
}

\frame{
\frametitle{``Conserved flux'' inflow boundary condition on $\widehat{f}_n$}
Non-standard choice of boundary condition: $\widehat{f}_n = \beta_n u_0$ on $\Gamma_{\rm in}$, induces smoother adjoint problems, \textcolor{red}{$\nor{\beta\cdot \grad v}_{L^2} = O(1)$}.
\begin{figure}[h!]
\centering
\subfigure[Primal problem]{\includegraphics[scale=.23]{figs/primal.png}}
\subfigure[Adjoint problem]{\includegraphics[scale=.23]{figs/adjoint.png}}
\caption{Under the new inflow condition, the wall-stop boundary condition is relaxed to a zero-stress condition at the outflow boundary of the adjoint problem.}
\end{figure}
}

\frame{
\frametitle{Test norms and adjoint solutions}
\textbf{Intuition:} the effectiveness of DPG under a test norm is governed by how a \textbf{specific test norm} measures the \textbf{solutions of the adjoint problem}.

\begin{figure}[!h]
\centering
\subfigure[Dirichlet inflow]{\includegraphics[scale=.35]{figs/dirichletBC.png}}
\subfigure[``Conserved flux'' inflow]{\includegraphics[scale=.35]{figs/newBC.png}}
\caption{DPG solutions to convection-diffusion for both inflow conditions using the $H^1$ test norm $\nor{\LRp{v,\tau}}_V \coloneqq \nor{v}_{H^1(\Omega)} + \nor{\tau}_{H^1(\Omega)}$.}
\end{figure}
}

%\frame{
%\frametitle{``Building blocks'' norm}
%Test norm quantities are robustly bounded from above by $\nor{u}_{L^2(\Omega)}$.  
%\[
%\|\left(v,\tau\right)\|_{V,K}^2 = \|v\|^2 + \epsilon \|\grad v\|^2 + \|\beta \cdot \grad v\|^2 + \| \div \tau\|^2 + \frac{1}{\epsilon}\|\tau\|^2.
%\]
%Problem: boundary layers in optimal test functions:
%\begin{figure}[!h]
%\centering
%\subfigure{\includegraphics[scale=.15]{figs/mesh1.png}}
%\subfigure{\includegraphics[scale=.15]{figs/sol1.png}}
%\caption{The $v$ component of the optimal test function corresponding to flux $\widehat{u} = x(1-x)$ on the bottom side of a unit element for $\epsilon = 0.01$.}
%\label{fig:boundaryTest}
%\end{figure}
%}
%\frame{
%\frametitle{Adjoint estimates, or ``norm building blocks''}
%For solutions $(v,\tau)$ of the adjoint equations, the following quantities are robustly bounded from above by $\nor{u}_{L^2(\Omega)}$.  
%\begin{align*}
%&\|v\|, \sqrt{\epsilon} \|\grad v\|, \|\beta \cdot \grad v\| \\
%&\| \div \tau\|, \frac{1}{\sqrt{\epsilon}}\|\tau\|.
%\end{align*}
%We will construct a test norm through a combination of the above quantities, such that
%\begin{itemize}
%\item $v$ and $\tau$ decoupled (no systems).
%\item Coefficients are of equal order after transforming to the unit element (no boundary layers).
%\end{itemize}
%}

\frame{
\frametitle{A mesh-scaled test norm}
Our test norm, as defined over a single element $K$, is now
\begin{align*}
\|\left(v,\tau\right)\|_{V,K}^2 = \|v\|^2 + \epsilon \|\grad v\|^2 + \|\beta \cdot \grad v\|^2 &\\
+\| \div \tau - \beta\cdot \grad v\|^2 + \min\left\{\frac{1}{\epsilon},\frac{1}{|K|}\right\}\|\tau\|^2&.
\end{align*}
which induces the proven \emph{robust} bound\footnote{\tiny\bibentry{DPGrobustness2}}\textsuperscript{,}\footnote{\tiny\bibentry{localConservationDPG}}
\[
\nor{u}_{L^2(\Omega)}+ \nor{\sigma}_{L^2(\Omega)} + \epsilon\nor{\widehat{u}} + \sqrt{\epsilon} \nor{\widehat{f}_n} \lesssim \nor{\LRp{u,\sigma,\widehat{u},\widehat{f}_n}}_E.
\]
}

\frame{
\frametitle{Erikkson-Johnson model problem}

On domain $\Omega = [0,1]^2$, with $\beta = (1,0)^T$, $ f = 0$ and boundary conditions
\begin{align*}
\widehat{\beta_nu - \sigma_n} &= \widehat{f}_n = u_0, \quad \beta_n \leq 0\\
\widehat{u} &= 0, \quad \beta_n > 0
\end{align*}

%Separation of variables gives an analytic solution. 
All numerical experiments done using Camellia codebase.\footnote{\tiny\bibentry{Camellia}}
%\[
%u(x,y) = C_0 + \sum_{n=1}^\infty C_n \frac{\exp(r_2(x-1)-\exp(r_1(x-1)))}{r_1\exp(-r2) - r_2\exp(-r1)}\cos(n\pi y)
%\]
\begin{figure}
\centering
\subfigure{
\includegraphics[width=1.4in]{figs/wallBC_exact_u.png}
}
\subfigure{
\includegraphics[width=1.4in]{figs/wallBC_exact_sigx.png}
}
\subfigure{
\includegraphics[width=1.4in]{figs/wallBC_exact_sigy.png}
}
\caption{Exact solution for $u$, $\sigma_x$, and $\sigma_y$ for $\epsilon = .01$, $C_1 = 1$, $C_n=0$, $n\neq 1$}
\end{figure}
}

\frame{
\frametitle{Error rates: numerical robustness}
\begin{figure}
\centering
\subfigure{
\includegraphics[width=2.275in]{figs/errorrates_wallBC.eps}
}
\subfigure{
\includegraphics[width=2.275in]{figs/L2energyratio_wallBC.eps}
}
%\subfigure{
%\includegraphics[scale=.22]{figs/u_pointdiff_wallBC.png}
%}
\caption{$L^2$/energy errors and their ratio for $\epsilon = .01, .001, .0001$.}
\end{figure}
}

\section{DPG for stationary nonlinear problems}
\begin{frame}{Outline}
  \tableofcontents[currentsection]
\end{frame}

\frame{
\frametitle{DPG for nonlinear problems}
Given some linearization technique (typically Newton-Raphson and pseudo-timestepping linearization), we measure 
\begin{itemize}
\item size of the linearized update $\Delta u$
\[
\|\Delta u\|_E \coloneqq \nor{B_u\Delta u}_{V'} = \nor{R_V^{-1} B_u\Delta u}_V 
\]
\item the nonlinear residual 
\[
\nor{R(u)}_E \coloneqq \nor{B(u) - \ell}_{V'}  = \nor{R_V^{-1}\LRp{B(u)-\ell}}_V 
\]
\end{itemize}
Preliminary experiments were done in 1D\footnote{\tiny\bibentry{DPGNS_1d}} and space-time.\footnote{\tiny\bibentry{DPGspacetime}}
}

\frame{
\frametitle{2D test case: Burgers equation}
\begin{columns}
\begin{column}{.45\textwidth}
\vspace{-.5cm}
\[
\pd{\left(u^2/2\right)}{x} + \pd{u}{y} + \epsilon \Delta u = f
\]
Burgers equation can be written with $\beta(u) = \LRp{u/2,1}$
\begin{align*}
\div\left(\beta(u)u-\sigma\right) &=f \\
\frac{1}{\epsilon}\sigma - \grad u &=0.
\end{align*}
i.e.\ nonlinear convection-diffusion on domain $[0,1]^2$.
\end{column}
\begin{column}{.52\textwidth}
\begin{figure}[!h]
\centering
\includegraphics[width = 2.15in]{figs/burgers.png}
\caption{Shock solution for a Burgers' equation with $\epsilon = 10^{-4}$ using Newton-Raphson.}
\end{figure}
\end{column}
\end{columns}
}
\frame{
Adaptivity begins with a cubic $4\times 4$ mesh.
\begin{figure}[!h]
\centering
\subfigure{\includegraphics[scale=.2]{figs/burgersMesh.png}}
\subfigure{\includegraphics[scale=.2]{figs/burgersZoom.png}}
\caption{Adaptive mesh after 9 refinements, and zoom view at point (.5,.5) with shock formation and $10^{-3}$ width line for reference.}
\end{figure}
}

\frame{
\frametitle{2D Compressible Navier-Stokes equations (ideal gas)}
Given density $\rho$, velocities $\mathbf{u} = \LRp{u_1,u_2}$ and temperature $T$,
\begin{align*}
\div \vecttwo{\rho u_1 }{\rho u_2} &= 0\\
\div \left(\vecttwo{\rho u_1^2+p }{\rho u_1 u_2} - \boldsymbol \sigma_{1}\right) &=0\\
\div \left(\vecttwo{\rho u_1 u_2}{\rho u_2^2+p } - \boldsymbol \sigma_{2}\right) &=0\\
\div \left(\vecttwo{((\rho e)+p)u_1}{((\rho e)+p)u_2} - \boldsymbol\sigma \mathbf{u} - \vec{q}\right) &=0\\
\frac{1}{2\mu} \boldsymbol \sigma - \frac{\lambda}{4\mu (\mu + \lambda)} { \rm tr}(\boldsymbol \sigma) \boldsymbol I &= \grad \mathbf{u} - \Reyn \, \boldsymbol \omega \\
\frac{1}{\kappa}\vec{q} &= \grad T
\end{align*}
}

\frame{
\frametitle{Extrapolation of test norms}
%\begin{columns}
%\begin{column}{.48\textwidth}
Convection-diffusion:
\begin{align*}
\div \left(\beta u - \sigma\right) &= f\\
\frac{1}{\epsilon}\sigma - \grad u &= 0.
\end{align*}
Navier-Stokes: defining vector variables $U = \{\rho,u_1,u_2,T\}$ and $\Sigma = \{{\boldsymbol \sigma},\mathbf{q},\omega\}$ and their linearized updates $\delta U, \delta \Sigma$,
\begin{align*}
\div \left(A_{\rm invisc}\delta U - A_{\rm visc}\delta\Sigma\right) &= R_{\rm conserv}(U,\Sigma)\\
E_{\rm visc} \delta\Sigma - \grad \delta U &= R_{\rm constit}(U,\Sigma)
\end{align*}
where $R_{\rm conserv}(U,\Sigma)$ and $R_{\rm constit}(U,\Sigma)$ are the conservation/constitutive residuals.
}

%\frame{
%\frametitle{Test norms over one element}
%Convection-diffusion:
%\begin{align*}
%\|\left(v,\tau\right)\|_{V,K}^2 =& \|v\|^2 + \|\beta \cdot \grad v\|^2 + \epsilon \|\grad v\|^2  \\
%&+\| \div \tau- \beta\cdot \grad v\|^2 + \min\left\{\frac{1}{\epsilon},\frac{1}{|K|}\right\}\|\tau\|^2.
%\end{align*}
%
%Navier-Stokes: let $V$ and $W$ be vectors of test functions $v_i$ and $\tau_i$.
%\begin{align*}
%\|\left(V,W\right)\|_{V,K}^2 =& \nor{V}^2  + \nor{A_{\rm invisc}^T \grad V}^2 + \frac{1}{\Reyn} \nor{A_{\rm visc}^T\grad V}^2 \\
%& + \nor{\div W-A_{\rm invisc}^T \grad V}^2 + \min\left\{1,\frac{1}{\Reyn|K|}\right\}\nor{E_{\rm visc}^TW}^2.
%\end{align*}
%}

\frame{
\frametitle{Carter's flat plate problem}
\begin{columns}
\begin{column}{.49\textwidth}
\begin{figure}[!h]
\centering
\includegraphics[scale=.32]{figs/flat_plate_BCs.pdf}
\caption{Carter flat plate problem on domain $[0,2]\times[0,1]$. Plate begins at $x=1$, ${\rm Re} = 1000$.}
\end{figure}
\end{column}
\begin{column}{.49\textwidth}
\begin{enumerate}
\item{} Symmetry boundary conditions.
\item{} Prescribed temperature and wall stagnation conditions.
\item{} Symmetry boundary conditions.
\item{} Inflow: conserved quantities specified using far-field values.
\item{} No outflow condition set.\footnotemark
\end{enumerate}
Stress/heat flux boundary conditions are set in terms of the momentum and energy fluxes. 
\end{column}
\end{columns}
\footnotetext{\tiny\bibentry{Shakib1991141}}
\let\thefootnote\relax\footnotetext{\tiny\bibentry{Carter}}
}

%\frame{
%\frametitle{Refinement level 0}
%\begin{figure}
%\centering
%\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/propMovie/u0.png}}
%\subfigure[$T$]{\includegraphics[scale = .23]{figs/propMovie/T0.png}}
%\subfigure{\includegraphics[scale = .35]{figs/propMovie/Mesh0.png}}
%\end{figure}
%}
%\setcounter{subfigure}{0}
%
%
%\frame{
%\frametitle{Refinement level 1}
%\begin{figure}
%\centering
%\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/propMovie/u1.png}}
%\subfigure[$T$]{\includegraphics[scale = .23]{figs/propMovie/T1.png}}
%\subfigure{\includegraphics[scale = .35]{figs/propMovie/Mesh1.png}}
%\end{figure}
%}
%\setcounter{subfigure}{0}
%
%
%\frame{
%\frametitle{Refinement level 2}
%\begin{figure}
%\centering
%\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/propMovie/u2.png}}
%\subfigure[$T$]{\includegraphics[scale = .23]{figs/propMovie/T2.png}}
%\subfigure{\includegraphics[scale = .35]{figs/propMovie/Mesh2.png}}
%\end{figure}
%}
%\setcounter{subfigure}{0}
%
%
%\frame{
%\frametitle{Refinement level 3}
%\begin{figure}
%\centering
%\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/propMovie/u3.png}}
%\subfigure[$T$]{\includegraphics[scale = .23]{figs/propMovie/T3.png}}
%\subfigure{\includegraphics[scale = .35]{figs/propMovie/Mesh3.png}}
%\end{figure}
%}
%\setcounter{subfigure}{0}
%
%
%
%\frame{
%\frametitle{Refinement level 4}
%\begin{figure}
%\centering
%\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/propMovie/u4.png}}
%\subfigure[$T$]{\includegraphics[scale = .23]{figs/propMovie/T4.png}}
%\subfigure{\includegraphics[scale = .35]{figs/propMovie/Mesh4.png}}
%\end{figure}
%}
%\setcounter{subfigure}{0}
%
%
%
%\frame{
%\frametitle{Refinement level 5}
%\begin{figure}
%\centering
%\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/propMovie/u5.png}}
%\subfigure[$T$]{\includegraphics[scale = .23]{figs/propMovie/T5.png}}
%\subfigure{\includegraphics[scale = .35]{figs/propMovie/Mesh5.png}}
%\end{figure}
%}
%\setcounter{subfigure}{0}
%
%\frame{
%\frametitle{Refinement level 6}
%\begin{figure}
%\centering
%\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/propMovie/u6.png}}
%\subfigure[$T$]{\includegraphics[scale = .23]{figs/propMovie/T6.png}}
%\subfigure{\includegraphics[scale = .35]{figs/propMovie/Mesh6.png}}
%\end{figure}
%}
%\setcounter{subfigure}{0}
%
%
%
%\frame{
%\frametitle{Refinement level 7}
%\begin{figure}
%\centering
%\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/propMovie/u7.png}}
%\subfigure[$T$]{\includegraphics[scale = .23]{figs/propMovie/T7.png}}
%\subfigure{\includegraphics[scale = .35]{figs/propMovie/Mesh7.png}}
%\end{figure}
%}
%\setcounter{subfigure}{0}
%
%
%\frame{
%\frametitle{Refinement level 8}
%\begin{figure}
%\centering
%\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/propMovie/u8.png}}
%\subfigure[$T$]{\includegraphics[scale = .23]{figs/propMovie/T8.png}}
%\subfigure{\includegraphics[scale = .35]{figs/propMovie/Mesh8.png}}
%\end{figure}
%}
%\setcounter{subfigure}{0}
%
%\frame{
%\frametitle{Refinement level 9}
%\begin{figure}
%\centering
%\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/propMovie/u9.png}}
%\subfigure[$T$]{\includegraphics[scale = .23]{figs/propMovie/T9.png}}
%\subfigure{\includegraphics[scale = .35]{figs/propMovie/Mesh9.png}}
%\end{figure}
%}
%\setcounter{subfigure}{0}
%
%
%\frame{
%\frametitle{Refinement level 10}
%\begin{figure}
%\centering
%\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/propMovie/u10.png}}
%\subfigure[$T$]{\includegraphics[scale = .23]{figs/propMovie/T10.png}}
%\subfigure{\includegraphics[scale = .35]{figs/propMovie/Mesh10.png}}
%\end{figure}
%}
%\setcounter{subfigure}{0}

\frame{
\frametitle{Reynolds number $\Reyn ={\rho_\infty V_\infty L_{\rm plate}}/{\mu_\infty} = 1000$, $p=2$}
\begin{figure}
\centering
\subfigure[$\rho$]{\includegraphics[width=2.25in]{figs/FlowFigs/PlateFigFiles/Re3/rho.png}}
\subfigure[$u_1$]{\includegraphics[width=2.25in]{figs/FlowFigs/PlateFigFiles/Re3/u1.png}}
\subfigure[$u_2$]{\includegraphics[width=2.25in]{figs/FlowFigs/PlateFigFiles/Re3/u2.png}}
\subfigure[$T$]{\includegraphics[width=2.25in]{figs/FlowFigs/PlateFigFiles/Re3/T.png}}
\end{figure}
}

\frame{
\frametitle{Reynolds number $1000$: adapted mesh}
\begin{figure}
\centering
\subfigure[Rescaled $\rho$]{\includegraphics[width=2.25in]{figs/FlowFigs/PlateFigFiles/Re3/rhoRescaled.png}}
\subfigure[Adapted mesh]{\includegraphics[width=2.25in]{figs/FlowFigs/PlateFigFiles/Re3/mesh12.png}}
\caption{Adapted $p=2$ mesh and rescaled density $\rho$ for Reynolds number $1000$.}
\end{figure}
}

\frame{
\frametitle{Reynolds number $1000$: solutions at stagnation point}
\begin{figure}
\centering
\subfigure[$\rho$]{\includegraphics[width=2in]{figs/FlowFigs/PlateFigFiles/Re3/rhoZoom.png}}
\subfigure[$u_1$]{\includegraphics[width=2in]{figs/FlowFigs/PlateFigFiles/Re3/u1Zoom.png}}
\subfigure[$u_2$]{\includegraphics[width=2in]{figs/FlowFigs/PlateFigFiles/Re3/u2Zoom.png}}
\subfigure[$T$]{\includegraphics[width=2in]{figs/FlowFigs/PlateFigFiles/Re3/TZoom.png}}
\end{figure}
}

\frame{
\frametitle{Reynolds number 10,000: 18 mesh refinements}
\begin{figure}[!h]
\centering
\subfigure[$\rho$]{\includegraphics[width=2.25in]{figs/FlowFigs/PlateFigFiles/Re4/rho.png}}
\subfigure[$u_1$]{\includegraphics[width=2.25in]{figs/FlowFigs/PlateFigFiles/Re4/u1.png}}
\subfigure[$u_2$]{\includegraphics[width=2.25in]{figs/FlowFigs/PlateFigFiles/Re4/u2.png}}
\subfigure[$T$]{\includegraphics[width=2.25in]{figs/FlowFigs/PlateFigFiles/Re4/T.png}}
\label{fig:Re10000}
\end{figure}
}

\frame{
\frametitle{Reynolds number 10,000: adapted mesh}
\begin{figure}
\centering
\subfigure[Rescaled $\rho$]{\includegraphics[width=2.25in]{figs/FlowFigs/PlateFigFiles/Re4/rhoRescaled.png}}
\subfigure[Adapted mesh]{\includegraphics[width=2.25in]{figs/FlowFigs/PlateFigFiles/Re4/mesh.png}}
\caption{Adapted mesh and rescaled density $\rho$ for Reynolds number 10,000.}
\end{figure}
}

\frame{
\frametitle{Reynolds number 10,000: solutions at stagnation point}
\begin{figure}
\centering
\subfigure[$\rho$]{\includegraphics[width=1.9in]{figs/FlowFigs/PlateFigFiles/Re4/rhoZoom.png}}
\subfigure[$u_1$]{\includegraphics[width=1.9in]{figs/FlowFigs/PlateFigFiles/Re4/u1Zoom.png}}
\subfigure[$u_2$]{\includegraphics[width=1.9in]{figs/FlowFigs/PlateFigFiles/Re4/u2Zoom.png}}
\subfigure[$T$]{\includegraphics[width=1.9in]{figs/FlowFigs/PlateFigFiles/Re4/TZoom.png}}
\end{figure}
}

\frame{
\frametitle{Solver upgrades for $\Reyn = $10,000}
\begin{itemize}
\item Static condensation: the DPG ultra-weak formulation yields 
\[
Ku = \arr{A}{B}{B^T}{D}\vecttwo{u_{\rm flux}}{u_{\rm field}} = \vecttwo{f}{g} = l
\]
where $D$ is block-diagonal. The system can be reduced to $\left(A-B D^{-1} B^T\right)u_{\rm flux} = f - B D^{-1} g$ with improved conditioning.\footnote{\tiny\bibentry{schurComplement}}
\item Anisotropic refinements using the error representation function\footnote{\tiny\bibentry{DPG3}}.
\item Line search to enforce positivity of $\rho$ and $T$.
%\[
%\rho \coloneqq \rho + \alpha \Delta \rho = \delta > 0, \qquad T \coloneqq T + \alpha \Delta T = \delta > 0
%\]
\item ``Effective'' CFL number: constraining $dt$ by $h_{\min}$.\footnote{\tiny\bibentry{Shakib1991141}}
\end{itemize}
}

\frame{
\frametitle{Holden ramp problem}
\begin{columns}
\begin{column}{.49\textwidth}
\begin{figure}[!h]
\centering
\includegraphics[width=2.25in]{figs/holden.eps}
\caption{Holden ramp problem on domain $[-1,2]\times[0,1]$. Plate begins at $x=-.5$.}
\end{figure}
\end{column}
\begin{column}{.49\textwidth}
\begin{itemize}
\item{} Prescribed temperature and wall stagnation conditions on plate and ramp.
\item{} Symmetry boundary conditions.
\item{} Inflow: conserved quantities specified using far-field values.
\item{} No outflow condition set.
\end{itemize}
\end{column}
\end{columns}
Stress/heat flux boundary conditions are set in terms of the momentum and energy fluxes. 
\let\thefootnote\relax\footnotetext{\tiny\bibentry{Carter}}
}

\frame{
\frametitle{Holden ramp problem - Mach 6, Reynolds 10,000}
\begin{figure}
\centering
\subfigure[$\rho$ rescaled]{\includegraphics[width=2.25in]{figs/FlowFigs/RampFigFiles/Ma6Re4/rhoRescaled.png}}
\subfigure[$u_1$]{\includegraphics[width=2.25in]{figs/FlowFigs/RampFigFiles/Ma6Re4/u1.png}}
\subfigure[$u_2$]{\includegraphics[width=2.25in]{figs/FlowFigs/RampFigFiles/Ma6Re4/u2.png}}
\subfigure[$T$]{\includegraphics[width=2.25in]{figs/FlowFigs/RampFigFiles/Ma6Re4/T.png}}
\caption{Solutions of Holden ramp problem for Mach 6 and Reynolds number 10,000 (approximately 34,000 with respect to plate length).}
\end{figure}
}

\frame{
\frametitle{Mach 6, Reynolds 10,000: solutions at shock}
\begin{figure}
\centering
\subfigure[Adapted mesh]{\includegraphics[width=3.25in]{figs/FlowFigs/RampFigFiles/Ma6Re4/mesh.png}}
\subfigure[Zoom of mesh at shock]{\includegraphics[width=2.25in]{figs/FlowFigs/RampFigFiles/Ma6Re4/shockMesh.png}}
%\subfigure[Zoom of $T$ at shock]{\includegraphics[width=2.25in]{figs/FlowFigs/RampFigFiles/Ma6Re4/Tshock.png}}
\subfigure[Zoom of rescaled $\rho$ at shock]{\includegraphics[width=2.25in]{figs/FlowFigs/RampFigFiles/Ma6Re4/rhoRescaledshock.png}}
\end{figure}
}

\frame{
\frametitle{Holden ramp problem - Mach 11.68, Reynolds 16,442.4}
\begin{figure}
\centering
\subfigure[$\rho$ rescaled]{\includegraphics[width=2.25in]{figs/FlowFigs/RampFigFiles/Ma11Re4/rhoRescaled.png}}
\subfigure[$u_1$]{\includegraphics[width=2.25in]{figs/FlowFigs/RampFigFiles/Ma11Re4/u1.png}}
\subfigure[$u_2$]{\includegraphics[width=2.25in]{figs/FlowFigs/RampFigFiles/Ma11Re4/u2.png}}
\subfigure[$T$]{\includegraphics[width=2.25in]{figs/FlowFigs/RampFigFiles/Ma11Re4/T.png}}
\caption{Solutions of Holden ramp problem for Mach 11.68 and Reynolds number 16,442.4 (approximately 55,800 with respect to plate length).}
\end{figure}
\let\thefootnote\relax\footnotetext{\tiny\bibentry{BenKirk}}
}

\frame{
\frametitle{Mach 11.68, Reynolds 16,442.4: solutions at shock}
\begin{figure}
\centering
\subfigure[Adapted mesh]{\includegraphics[width=3.25in]{figs/FlowFigs/RampFigFiles/Ma11Re4/mesh.png}}
\subfigure[Zoom of mesh at shock]{\includegraphics[width=2.25in]{figs/FlowFigs/RampFigFiles/Ma11Re4/shockMesh.png}}
%\subfigure[Zoom of $T$ at shock]{\includegraphics[width=2.25in]{figs/FlowFigs/RampFigFiles/Ma11Re4/Tshock.png}}
\subfigure[Zoom of rescaled $\rho$ at shock]{\includegraphics[width=2.25in]{figs/FlowFigs/RampFigFiles/Ma11Re4/rhoshock.png}}
\end{figure}
}

\frame{
\frametitle{Normal heat flux for Holden flow}
\begin{figure}
\centering
\subfigure[Mach 6, Reynolds 10,000]{\includegraphics[width=2.25in]{figs/FlowFigs/RampFigFiles/Ma6Re4/heatFlux.png}}
\subfigure[Mach 11.68, Reynolds 16,442.4]{\includegraphics[width=2.25in]{figs/FlowFigs/RampFigFiles/Ma11Re4/heatFlux.png}}
\caption{Normal heat flux $q_n$ over plate and ramp for the Holden problem (\textcolor{blue}{blue} is heat flux over the plate, while \textcolor{brown}{brown} is heat flux over the ramp, plotted against the $x$-coordinate along the ramp).}
\end{figure}
}

\frame{
\frametitle{Summary of work}
\begin{itemize}
\item \textbf{Area A:}
\begin{itemize}
\item Developed a provably robustness DPG method and theory on global properties of DPG test spaces for convection-diffusion problems.
\item Developed theory for extension of DPG to nonlinear problems.  
\end{itemize}
\item \textbf{Area B:}
\begin{itemize}
\item Contributions to parallel $hp$-adaptive Camellia\footnote{\tiny\bibentry{Camellia}} (anisotropic refinements, Schur-complement linear solver).
\item Construction of a higher order adaptive Navier-Stokes solver.
\end{itemize}
\item \textbf{Area C:}
\begin{itemize}
\item Solution of 2D convection-diffusion and viscous Burgers' equation.
\item Solution of 2D compressible Navier-Stokes equations for the Carter flat plate and Holden compression ramp problems.
\end{itemize}
\end{itemize}
}

\frame{
\frametitle{Future directions}
\begin{itemize}
\item Navier-Stokes: non-primitive variables, artificial diffusion methods.
\item Scalability: GPU architectures, preconditioners and multigrid solvers. 
\item Alternative discretizations: saddle point formulation\footnote{\tiny\bibentry{H1DPG}}
\begin{figure}
\centering
\subfigure[Problem setup]{\includegraphics[width=1.3in]{figs/discForcing.eps}}
\subfigure[$p=2$ mesh]{\includegraphics[width=1.3in]{figs/discontinuousForcing.png}}
\subfigure[Solution]{\includegraphics[width=1.3in]{figs/discontinuousForcingMesh.png}}
\caption{Results using FEniCS for convection-diffusion under a $C_0$ basis.}
\end{figure}
\end{itemize}
}

\frame{
\begin{center}
Thank you!\\
\vspace{.5cm}
Questions?
\end{center}
\begin{figure}
\centering
\subfigure[3D slice]{\includegraphics[width=1.75in]{figs/3DSlice.png}}
\subfigure[3D mesh slice]{\includegraphics[width=1.75in]{figs/3DMeshSlice.png}}
\caption{An adaptive convection-diffusion experiment in FEniCS with $\epsilon = 10^{-6}$ and a $C_0$ basis.}
\end{figure}
}

\frame{
\frametitle{Extended slides: Oscillations for nonlinear shock problems}
1D linearized Burgers': $\LRp{u\Delta  u}_{,x} - \epsilon \Delta u_{,xx} = -r(u) = -\LRp{\LRp{\frac{u^2}{2}}_{,x} - \epsilon u_{,xx}}$
%: $\LRp{\frac{u^2}{2}}_{,x} - \epsilon u_{,xx} = f$.   is
\begin{figure}
\centering
\subfigure[$u_{\epsilon}(x)$ for $\epsilon = .01$]{\includegraphics[width=1.5in]{figs/Burgers1DPlots/u.eps}}
\subfigure[$r(u_{2\epsilon})$ for $\epsilon = .01$]{\includegraphics[width=1.5in]{figs/Burgers1DPlots/residual100.eps}}
\subfigure[$r(u_{2\epsilon})$ for $\epsilon = .001$]{\includegraphics[width=1.5in]{figs/Burgers1DPlots/residual1000.eps}}
\caption{Nonlinear residual/forcing term $r(u_{2\epsilon})$.  Linearization is done around the exact solution for $2\epsilon$.}
\end{figure}
}

\frame{
\frametitle{Extended slides: Navier-Stokes stress law}
$\boldsymbol\sigma$ for a Newtonian fluid: $\sigma_{ij} = \mu(u_{i,j} + u_{j,i}) + \lambda u_{k,k} \delta_{ij}$.
%We can invert the stress tensor under isotropic and plane strain assumptions to get
\[
\frac{1}{2}\left(\grad  U + \grad ^T  U\right) = \frac{1}{2\mu} \sigma_{ij} - \frac{\lambda}{4\mu (\mu + \lambda)} \sigma_{kk}\delta_{ij}
\]
The symmetric part of the gradient is
\[
\frac{1}{2}\left(\grad  U + \grad ^T  U\right) = \grad  U - \boldsymbol \omega
\]
%where $\boldsymbol \omega$ is the antisymmetric part of the infinitesimal strain tensor:
%\[
%\boldsymbol \omega = \frac{1}{2}\left(\grad  U - \grad ^T  U\right).
%\]
Our final form is
\begin{align*}
\grad  U - \boldsymbol \omega = \frac{1}{2\mu} \boldsymbol \sigma - \frac{\lambda}{4\mu (\mu + \lambda)} { \rm tr}(\boldsymbol \sigma) \boldsymbol I.
\end{align*}
By enforcing strong symmetry of $\boldsymbol \sigma$, taking the antisymmetric part implicitly defines $\boldsymbol \omega$ to be the antisymmetric part of $\grad u$. 
}

%\frame{
%\frametitle{Extended literature review}
%\begin{itemize}
%\item Finite difference/volume methods, upwind stencils, monotonicity, ENO/WENO. 
%\item Artificial diffusion, flux limiters/discontinuity-capturing operators.
%\item DG fluxes can be interpreted as residual-based stabilizations similar to SUPG.\footnote{\tiny\bibentry{Brezzi20063293}}
%\item Hybridized DG (HDG) methods introduce independent unknowns for numerical traces. Fluxes are defined in terms of DG fluxes and the numerical trace.\footnote{\tiny\bibentry{hybridDG}}
%\end{itemize}
%}

%\frame{
%\frametitle{Specifics of the ultra-weak formulation}
%
%Given a first order system $Au = f$, multiply by test function $v$ and integrate 
%\[
%\LRp{Au,v} = \LRa{\gamma\LRp{Au},v} + \LRp{u,A^*_hv} = \LRp{f,v}
%\]
%If $u \in L^2\LRp{\Omega}$, the trace of $u$ is undefined, so we identify boundary terms $\LRa{\gamma\LRp{Au},v}_{\Gh} = \LRa{\widehat{u},v}_{\Gh}$ as unknowns $\widehat{u}$ on $\Gh$. For convection-diffusion,
%\begin{align*}
%b\left(\left(u,\sigma, \widehat{u}, \widehat{f}_n\right),
%\left( v, \tau \right)\right) = \left(u,\grad_h\cdot \tau - \beta \cdot \grad_h
%v\right)_{\Oh} + \left(\sigma, \epsilon^{-1} \tau + \grad_h v\right)_{\Oh}&\\
%  - \LRa{\jump{\tau\cdot n}, \widehat{u} }_{\Gh} + \LRa{\widehat{f}_n,\jump{v} }_{\Gh}&,
%\end{align*}
%where 
%\begin{align*}
%\widehat{f}_n \coloneqq \beta_n u - \sigma_n \in H^{-1/2}(\Gh), \quad \widehat{u} \in H^{1/2}(\Gh)
%\end{align*}
%with the minimum energy extension norm on $\Gh$. We note that $H^{\pm 1/2}(\Gh)$ are \emph{closed} subspaces of $\prod_K H^{\pm 1/2}(\partial K)$. 
%}

\bibliographystyle{plain}
{\scriptsize\bibliography{paper,NSNotes,DPGarticles}
}

\end{document}